Sketch Recognition using Particle Swarm Optimization Algorithm

### 2009-04-12 trunk/draft/System.tex
after grammer correct now will print to review

 In \textbf{Feature Extraction} step, a feature vector consists of a set of statistical, geometrical and spatial features is extracted from the segmented strokes. The feature vector is used as input for the next step.   %the system computes composite set of features some are statistical other are spatial features based on the type primitives.% The system extract segment based and statistical based features from the set of segments that the clustering algorithms produce.

@@

\subsection{Critical Point Detection}
\label{sec:CriticalPointDetection}
After computing speed, time difference and curvature information the system proceeds to detect the points with low velocity and high curvatures. Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated. Then a threshold \textit{th} is used to separate the curve into regions; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and the dominate points $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.  % (as shown are redundant)
\begin{figure}
	\centering
		\includegraphics[scale=0.5]{images/speedreg.eps}
	\caption{Possible Dominant Points Calculations}
	\label{fig:MaxRegioi}
\end{figure}

@@

The system repeats this process to curvature, direction, time difference and speed curves. For each curve the points labeled as possible dominant points $P_{pd}$ are saved into a single array. The points that are repeatedly labeled are given a higher score than the other points. This score is then used to sort the  $P_{pd}$ points array. The array that contains possible points $P_{pd}$ is used as input for the next segmentation step. Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate points $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.

\begin{figure}
	\centering
		\includegraphics{images/ppd.eps}
	\caption{Posible domninate points}% a) Possible dominate point b) Particle encoding  
	\label{fig:LabelsPPD}
\end{figure}

@@

The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set of dominate points that defines the stroke with minimum segmentation error. In other words, the system tries to find the fewer number of points (1's) location in the particle array which gives minimum segmentation error.

@@

\subsubsection{Polygon Division Algorithm \textsl{AlgS1}}
\label{sec:PolygonDivisionAlgorithm}
The first algorithm, the algorithm tries to segment the stroke as a polygon. The final solution is a set of line segments that best define the input stroke. Given the inputs stroke let's define the set of points in the stroke as  $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ which is the set of consecutive points from start of stroke till the end. The arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The line
$\overline{x_i x_j} $ is the straight line connecting point $x_i$ to point $x_j$. The approximation error is computed by the equation \ref{eq:ErrorSwarm1} 
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation}
 where $M$ is the number of dominant points in this solution.  The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$ \cite{PolygonApproximationPSO}. Figure \ref{fig:DPSOERROR} shows a graphical representation of the error. The fitness is computed using equation \ref{eq:fitnessSwarm1} 
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation}%\]
where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as $P_{pd}$  and $E$ is the computed error and $\varepsilon$ is the error threshold.  If the solution produced an error that exceed the error threshold $\varepsilon$ the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices'.  This fitness function optimize two goals the first is to move the solution into the feasible solution space with acceptable error bounds and second is to fly the particle to a new position which may result in a polygon with a fewer number of vertices'\cite{PolygonApproximationPSO}. 
\begin{figure}
	\centering
  \includegraphics[scale=0.8]{images/pso1.eps}			
	\caption{ALS1 Error}% a) Possible dominate point b) Particle encoding  
	\label{fig:DPSOERROR}
\end{figure}
 %The error is chosen to in favor of larger than the threshold is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices'.\\ %we will want to lower the number of vertices'. \\
%if we say that s
%Alg1:
%Alg2: 
\subsubsection{Curve Divisions Algorithm \textsl{AlgS2}}
\label{sec:PolygonDivisionAlgorithm}

@@

\subsubsection{Solution Refine procedures } 
%\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
 After each loop in the swarm algorithm, each particle loops on the set of selected dominant points to enhance the solution. Each dominant point is checked to find if it was labeled before as a possible dominant (computed as in section \ref{sec:Preprocessing}), if not the point is moved to the nearest labeled point.
   After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure. For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{sec:Preprocessing}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$. After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$. If two points are nearer than $min_D$\footnote{Constant selected to be 10 pixels} then one of the points is removed.

\subsection{Choosing the best fit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. Equation \ref{eq:bestFit1} and \ref{eq:bestFit2} are used to evaluate each segmentation.

\begin{equation}
\label{eq:bestFit1}
E_{area}=A_{st}-A_{sg}
 \end{equation}
 \begin{equation}
\label{eq:bestFit2}
E_{tot}=E_{area}+E_{alg}
 \end{equation}
 where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmenation, $E_{alg}$ is the approximation error of the segmentation. The segmentation which crosspond to the minimum $E_{tot}$ is the best segmentation.

\section{Hybird Segmentation Algorithm}
\label{sec:BenchMarckAlgorithm}

@@

%\section{Stroke Clustring Algoirthm }
%\label{sec:ClustringAlgoirthm}
%
%When the user finishes drawing a stroke the segmentation algorithm generate a list of segments that represent this stroke. The system then groups the segments generated from the segmentation algorithm. %These segments are grouped into a symbol and tested if they can be classified into one of the known symbols. If no identification is achieved the segments are added into a list of unclassified segments. When a user draws another stroke the segmented stroke is added to the list and the process is repeated until all the segments are classified to a known symbol.\\
%
%After the user draw all strokes of the symbol he has to wait 10 seconds or press finish button beside the drawing area. The set of unrecognized strokes is grouped together along with their segmentation as input to the feature extraction process.

@@

\begin{description}
 \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
\begin{description}
	 \item[Segments] Number of segments in the symbol.
	 \item [Strokes] Number of strokes or partical strokes that created the symbols.  
		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item [Curves] Number of curves or ellipses in the symbol. 
		\item [Lines] Number of lines in the symbol. 
		\item [Perpendicular lines] Number of perpendicular lines. 
		\item [Parallel lines] Number of parallel lines. 
		\item [Intersections] Number of intersection between lines and curves. 
		\item [T intersections] Number of T intersections. 
		\item [L intersections] Number of L intersections. 
		\item [X intersections] Number of X intersections.
		\item [Size Ratio] Ratio between width to height of the symbol.

@@

\end{description}
\item [Rubine Feature Set (FS2)] Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box.It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It  gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item  Total time of stroke 
\end{enumerate}

>>>>

 In \textbf{Feature Extraction} step, a feature vector consists of a set of statistical, geometrical and spatial features are extracted from the segmented strokes. The feature vector is used as input for the next step.   %the system computes composite set of features some are statistical other are spatial features based on the type primitives.% The system extract segment based and statistical based features from the set of segments that the clustering algorithms produce.

@@

\subsection{Critical Point Detection}
\label{sec:CriticalPointDetection}
After computing speed, time difference and curvature information the system proceeds to detect the points with low velocity and high curvatures. Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated. Then a threshold \textit{th} is used to separate the curve into regions; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and the dominant points $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.  % (as shown are redundant)
\begin{figure}
	\centering
		\includegraphics[scale=0.5]{images/speedreg.eps}
	\caption{Possible Dominant Points Calculations}
	\label{fig:MaxRegioi}
\end{figure}

@@

The system repeats this process to curvature, direction, time difference and speed curves. For each curve the points labeled as possible dominant points $P_{pd}$ are saved into a single array. The points that are repeatedly labeled are given a higher score than the other points. This score is then used to sort the  $P_{pd}$ points array. The array that contains possible points $P_{pd}$ is used as input for the next segmentation step. Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant points $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.

\begin{figure}
	\centering
		\includegraphics{images/ppd.eps}
	\caption{Possible dominant points}% a) Possible dominant point b) Particle encoding  
	\label{fig:LabelsPPD}
\end{figure}

@@

The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set of dominant points that defines the stroke with minimum segmentation error. In other words, the system tries to find the fewer number of points (1's) location in the particle array which gives minimum segmentation error.

@@

\subsubsection{Polygon Division Algorithm \textsl{AlgS1}}
\label{sec:PolygonDivisionAlgorithm}
The first algorithm, the algorithm tries to segment the stroke as a polygon. The final solution is a set of line segments that best define the input stroke. Given the inputs stroke let's define the set of points in the stroke as  $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ which is the set of consecutive points from start of stroke till the end. The arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The line
$\overline{x_i x_j} $ is the straight line connecting point $x_i$ to point $x_j$. The approximation error is computed by the equation \ref{eq:ErrorSwarm1} 
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation}
 where $M$ is the number of dominant points in this solution.  The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$ \cite{PolygonApproximationPSO}. Figure \ref{fig:DPSOERROR} shows a graphical representation of the error. The fitness is computed using equation \ref{eq:fitnessSwarm1} 
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation}%\]
where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as $P_{pd}$  and $E$ is the computed error and $\varepsilon$ is the error threshold.  If the solution produced an error that exceed the error threshold $\varepsilon$ the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices'.  This fitness function optimize two goals the first is to move the solution into the feasible solution space with acceptable error bounds and second is to fly the particle to a new position which may result in a polygon with a fewer number of vertices'\cite{PolygonApproximationPSO}. 
\begin{figure}
	\centering
  \includegraphics[scale=0.8]{images/pso1.eps}			
	\caption{AlgS1 Error}% a) Possible dominant point b) Particle encoding  
	\label{fig:DPSOERROR}
\end{figure}
 %The error is chosen to in favor of larger than the threshold is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices'.\\ %we will want to lower the number of vertices'. \\
%if we say that s
%Alg1:
%Alg2: 
\subsubsection{Curve Divisions Algorithm \textsl{AlgS2}}
\label{sec:PolygonDivisionAlgorithm}

@@

\subsubsection{Solution Refine procedures} 
%\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
 After each loop in the swarm algorithm, each particle loops on the set of selected dominant points to enhance the solution. Each dominant point is checked to find if it was labeled before as a possible dominant (computed as in section \ref{sec:Preprocessing}), if not the point is moved to the nearest labeled point.
   After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure. For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{sec:Preprocessing}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. After that the particles are tested to make sure that the distance between every two successive dominant point is larger than the constant $min_D$. If two points are nearer than $min_D$\footnote{Constant selected to be 10 pixels} then one of the points is removed.

\subsection{Choosing the best fit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. The segmentation error($E_{alg}$ in Equation \ref{eq:bestFit2})represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line. In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}. The additional computation is necessary to  compare the two segmentation fairly. 
\begin{equation}
\label{eq:bestFit1}
E_{area}=A_{st}-A_{sg}
 \end{equation}
 \begin{equation}
\label{eq:bestFit2}
E_{tot}=E_{area}+E_{alg}
 \end{equation}
 where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error. The segmentation which cross ponds to the minimum $E_{tot}$ is the best segmentation.

\section{Hybrid Segmentation Algorithm \textsl{Alg3}}
\label{sec:BenchMarckAlgorithm}

@@

%\section{Stroke Clustring Algorithms }
%\label{sec:ClustringAlgoirthm}
%
%When the user finishes drawing a stroke the segmentation algorithm generate a list of segments that represent this stroke. The system then groups the segments generated from the segmentation algorithm. %These segments are grouped into a symbol and tested if they can be classified into one of the known symbols. If no identification is achieved the segments are added into a list of unclassified segments. When a user draws another stroke the segmented stroke is added to the list and the process is repeated until all the segments are classified to a known symbol.\\
%
%After the user draw all strokes of the symbol he has to wait 10 seconds or press finish button beside the drawing area. The set of unrecognized strokes is grouped together along with their segmentation as input to the feature extraction process.

@@

\begin{description}
 \item [Structural and geometrical Features (FS1)] Features defines the structure of the geometrical symbol.  
\begin{description}
	 \item[Segments] Number of segments in the symbol.
	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item [Curves] Number of curves or ellipses in the symbol. 
		\item [Lines] Number of lines in the symbol. 
		\item [Perpendicular lines] Number of perpendicular lines. 
		\item [Parallel lines] Number of parallel lines. 
		\item [Intersections] Number of intersection between lines and curves. 
		\item [T intersections] Number of T intersections. 
		\item [L intersections] Number of L intersections. 
		\item [X intersections] Number of X intersections.
		\item [Size Ratio] Ratio between width to height of the symbol.

@@

\end{description}
\item [Rubine Feature Set (FS2)] Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item Total time of stroke 
\end{enumerate}

### 2009-04-12 trunk/draft/chapterSurvey.tex
after grammer correct now will print to review

% i need to write a history then small introudction to the survey...
%  i have to state that the next  sections will generally group segmentation and symbol recognition task. 
\section{Preprocessing and Segmentation}
\label{sec:preprocessing}
%this section should describe how others did the first preprocessing of the problem. i have to summarize what i wrote before. 
 % first paragrah about how stroke is a messing thing and naturally it is ambigues. 
  In a system that supports free hand sketching, sketching should be as natural as possible. This means that the user can draw symbols using single or multiple strokes without constrains. The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper. But sketches are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users. This led researchers to perform sort of preprocessing before starting the sketch recognition task itself. The next sections explain different preprocessing techniques.

@@

One of the first systems in gestures recognition is Rubine's algorithm \cite{gestureexample12} . He used Gaussian statistical classifier to classify gestures.  Simple features as length of stroke, smoothness of stroke, the angle that the first part of the gesture is drawn at and properties of the bounding box of the stroke are used to train the classifier to recognize  symbols. In our system, we combine Rubine features with others to include the information on how the user draws sketch while recognizing the symbols. In \cite{Long00}  Long shows how similarities between gestures can be analyzed.  He used this information to identify gestures that are likely to confuse the recognition system. This helped interface designers to choose gestures that are easily recognized.

@@

 Avoiding graph matching problems, research tried to use languages that describe shapes instead of graphs. Visual descriptive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives. Figures \ref{fig:arrow1} and \ref{fig:arr2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}. Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.

 Several visual grammar were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}. \cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified. The system  automatically generates Java code for the recognizer for each object. The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.  These systems are less computational expensive than graph matching approach but they also operate based on the output of the segmentation process. Any error in the segmentation process lead to huge error while parsing the whole sketch. Moreover, symbols cannot share strokes between them which influence and restrict how the user draw symbols. 
 % complexity of the graph matching problem 
\begin{figure}
	\centering
		\subfigure[An Arrow]
			{\label{fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
			\hfill
		\subfigure[Arrow grammar definition]
		 {\label{fig:arr2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
	\caption[Arrow definition]{}
		\label{fig:arrowladderdef1}
\end{figure}

@@

drawing back from properties of individual strokes several researcher tried to use shape properties on the whole shape to classify shapes. They used global features that tries to define the entire shape properties.

 Several systems \cite{DiagramOfflineConvexHull,MulitStrokeConvexHull} presented features like ratio of bounding box to convex hull area and the ratio of the perimeter to the area.  These global feature can be used to distinguish between classes. For example, a rectangle can be distinguish from a triangle by looking to the ratio of area of bounding box to the area of the convex hull. The triangle ratio will be much lower than that of the rectangle.

@@

 Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tend to have. These properties dose not depend on the order or number of strokes and thus are only based on appearance of the shape not how the user draw it. However, individual details of the shape are impossible to distinguish using these properties. For example, it is impossible to use  to  differentiate between a current source symbol (circle containing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a tilate Figure \ref{fig:acSource}).  Finer details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distinguished using this method. These symbols requires a system to represent different levels of shape details.

@@

In the third approach, the recognition focus on appearance of shapes contrary to individual strokes or global shape properties. Kara \cite{imagetrainable48} used this approach while presenting a system that matches the input shape into a one of the  prototypes in the database. Figure \ref{fig:template} shows an example of prototypes of some of the shapes in the system. Input shapes are first pose and scale normalized then four image similarity measure are used to match input shapes to the database prototypes. He used two measures (Hausdor distance)  to measure similarity and two coefficients (Yule, Tanimoto) to measure dissimilarity. His approach allows the system to recognized a shape using only a single training sample as it is based on the shape prototype. However, the shape prototype must represent all  transformation and variation the input shape can have. Moreover, they restrict the image representation to 48x48 pixel due to computational reasons. This restriction leads to eliminating some of the fine details that may be needed to distinguish some shape classes. There results on isolated symbols seems promising but the system have not beet tested on full sketch that tends to have high overlapping between symbols.

@@

In all of these system only \cite{SketchRead2007} using stroke based recognition have been implemented to the fully creat a system that free draw sketch and understand them. This is because the emphasis on the interactive side of the sketch sytem which result in large difficulties in processing free hand sketches. Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be grouped. These systems assume that users will finish drawing a shape before moving to the next one \cite{multistrokeEvaluation,Cali63,geometrydomain49}. This let them to temporally group strokes and then recognize them based on their strokes or appearance properties. \cite{AlvaradoDigital} argued that this assumption dose not hold in natural sketches.

Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}. The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process. Natural sketch are difficult to segment and group to create reliable fragment of geometeric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols.  Since these systems are dependable on the segmentation results, they either performe recognition using undependable segmentation results or consider different ways to segment the input sketch.

To avoid most problems in such system, our system uses hybrid of both global shape and hierarchical description methods. Currently, our system recognize on symbols at a time. Each symbol is segmented and hierarchical features are computed.  Features  that are represent the global description of the shape are also generated and appended to the hierarchical features. The recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.

@@

Modeling how a biological swarm has been recently studied by biologist and computer scientists. Scientist in the field or artificial life need to understand how such "social animals" interact, achieve goals and evolve. Moreover, swarm behaviors that achieves some kind of swarm intelligence is increasingly interesting for engineers as it they can be applied in optimization. Looking from a higher view, agents in a swarm are cooperating to achieve some purposeful behavior and achieve some common goal. For example, ant groups looking for food source. Large groups of simple independent agents seems to generate this behavior in what is called 'collective intelligence'. Each agent has a set of simple rules to follow, and according to the interaction of the entire group the objective is met\cite{swarmGeneral}. %with other neighbourhood agents.

Reynolds \cite{swarmflock} created 'boid' which simulated the motion of a flock of birds. Each boid is a simple agent that navigate according to his perception and a set of basic rules.  Kube and Zhange \cite{antsimulation} introduced a simulation model for ants behavior and used in cooperative transportation. Ants leaves a trail of its movement, so when an ant find food it cannot carry alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.   Kube and Zhange \cite{antsimulation,swarmGeneral} system seems to correctly models ants behavior with a methods of task modeling.

@@

ToADD

>>>>

% i need to write a history then small introduction to the survey...
%  i have to state that the next  sections will generally group segmentation and symbol recognition task. 
\section{Preprocessing and Segmentation}
\label{sec:preprocessing}
  In a system that supports free hand sketching, sketching should be as natural as possible. This means that the user can draw symbols using single or multiple strokes without constrains. The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper. But sketches are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users. This led researchers to perform sort of preprocessing before starting the sketch recognition task itself. The next sections explain different preprocessing techniques.

@@

One of the first systems in gestures recognition is Rubine's algorithm \cite{gestureexample12} . He used Gaussian statistical classifier to classify gestures.  Simple features as length of stroke, smoothness of stroke, the angle that the first part of the gesture is drawn at and properties of the bounding box of the stroke are used to train the classifier to recognize symbols. In our system, we combine Rubine features with others to include the information on how the user draws sketch while recognizing the symbols. In \cite{Long00}, Long shows how similarities between gestures can be analyzed.  He used this information to identify gestures that are likely to confuse the recognition system. This helped interface designers to choose gestures that are easily recognized.

@@

 Avoiding graph matching problems, research tried to use languages that describe shapes instead of graphs. Visual descriptive languages represent the shape into tokens of geometric primitives or relations between smaller geometric primitives. Figures \ref{fig:arrow1} and \ref{fig:arr2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}. Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.

 Several visual grammar were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}. \cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified. The system automatically generates Java code for the recognizer for each object. The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.  These systems are less computational expensive than graph matching approach but they also operate based on the output of the segmentation process. Any error in the segmentation process leads to huge error while parsing the whole sketch. Moreover, symbols cannot share strokes between them which influence and restrict how the user draws symbols. 
 % complexity of the graph matching problem 
\begin{figure}
	\centering
		\subfigure[An Arrow]
			{\label{fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
			\hfill
		\subfigure[Arrow grammar definition]
		 {\label{fig:arr2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
	\caption[Arrow definition]{}
		\label{fig:arrowladderdef1}
\end{figure}

@@

Drawing back from properties of individual strokes several researcher tried to use shape properties on the whole shape to classify shapes. They used global features that try to define the entire shape properties.

 Several systems \cite{DiagramOfflineConvexHull,MulitStrokeConvexHull} presented features like ratio of bounding box to convex hull area and the ratio of the perimeter to the area.  These global features can be used to distinguish between classes. For example, a rectangle can be distinguished from a triangle by looking to the ratio of area of bounding box to the area of the convex hull. The triangle ratio will be much lower than that of the rectangle.

@@

 Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tends to have. These properties do not depend on the order or number of strokes and thus are only based on appearance of the shape not how the user draws it. However, individual details of the shape are impossible to distinguish using these properties. For example, it is impossible to use to differentiate between a current source symbol (circle containing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a wave Figure \ref{fig:acSource}).  Finer details of the symbols are not represented in this approach, thus shapes that have small's details or only significant conceptual variation are not distinguished using this method. These symbols require a system to represent different levels of shape details.

@@

In the third approach, the recognition focuses on appearance of shapes contrary to individual strokes or global shape properties. Kara \cite{imagetrainable48} used this approach while presenting a system that matches the input shape into a one of the prototypes in the database. Figure \ref{fig:template} shows an example of prototypes of some of the shapes in the system. Input shapes are first pose and scale normalized then four image similarity measures are used to match input shapes to the database prototypes. He used two measures (Hausdor distance) to measure similarity and two coefficients (Yule, Tanimoto) to measure dissimilarity. His approach allows the system to recognize a shape using only a single training sample as it is based on the shape prototype. However, the shape prototype must represent all transformation and variation the input shape can have. Moreover, they restrict the image representation to 48x48 pixels due to computational reasons. This restriction leads to eliminating some of the fine details that may be needed to distinguish some shape classes. There results on isolated symbols seem promising but the system has not been tested on full sketch that tends to have high overlapping between symbols.

@@

In all of these system only \cite{SketchRead2007} using stroke based recognition have been implemented to the fully create a system that free draw sketch and understand them. This is because the emphasis on the interactive side of the sketch system which result in large difficulties in processing free hand sketches. Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be grouped. These systems assume that users will finish drawing a shape before moving to the next one \cite{multistrokeEvaluation,Cali63,geometrydomain49}. This let them to temporally group strokes and then recognizes them based on their strokes or appearance properties. \cite{AlvaradoDigital} argued that this assumption does not hold in natural sketches.

Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}. The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process. Natural sketch are difficult to segment and group to create reliable fragment of geometric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols.  Since these systems are dependable on the segmentation results, they either perform recognition using undependable segmentation results or consider different ways to segment the input sketch.

To avoid most problems in such system, our system uses hybrid of both global shape and hierarchical description methods. Currently, our system recognize on symbols at a time. Each symbol is segmented and hierarchical features are computed.  Features that are representing the global description of the shape are also generated and appended to the hierarchical features. The recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.

@@

Modeling how a biological swarm has been recently studied by biologist and computer scientists. Scientist in the field or artificial life need to understand how such "social animals" interact, achieve goals and evolve. Moreover, a swarm behavior that achieves some kind of swarm intelligence is increasingly interesting for engineers as it they can be applied in optimization. Looking from a higher view, agents in a swarm are cooperating to achieve some purposeful behavior and achieve some common goal. A good example of this cooperation is ant groups looking for food source. A large group of simple independent agents seems to generate this behavior in what is called 'collective intelligence'. Each agent has a set of simple rules to follow, and according to the interaction of the entire group the objective is met\cite{swarmGeneral}. %with other neighbourhood agents.

Reynolds \cite{swarmflock} created 'boid' which simulated the motion of a flock of birds. Each boid is a simple agent that navigates according to his perception and a set of basic rules.  Kube and Zhange \cite{antsimulation} introduced a simulation model for ant's behavior and used in cooperative transportation. Ants leaves a trail of its movement, so when an ant find food it cannot carry alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.   Kube and Zhange \cite{antsimulation,swarmGeneral} system seems to correctly models ants behavior with a methods of task modeling.

@@
